### 2017-07-08 article/paper.tex
Application Chapter finished

\documentclass{llncs}
\usepackage{makeidx}
\usepackage[utf8]{inputenc}
%\usepackage{graphics}
%\usepackage[pdftex]{graphicx}
\usepackage{listings}
\usepackage{float}
\usepackage{graphicx}

\begin{document}

\title{Waypoint navigation for an airplane UAV using MAVlink in JdeRobot framework}
\author{José Antonio Fernández\inst{1} \and Jose María Cañas Plaza\inst{2}}
\institute{Universidad Rey Juan Carlos}
\titlerunning{Article Abv}
\maketitle

\begin{abstract}
In order to make accesible to \texttt{JdeRobot} the command sets of \texttt{MAVLink} a driver is needed. The \texttt{APM Server driver} ``translates'' the \texttt{JdeRobot} interfaces objects in \texttt{MAVLink's} commands an viceversa, giving access to sensors and actuators of the robot. With the typical \texttt{JdeRobot} architecture applications can run in different machines (not necessarily aboard) and can be writen in different programming languages. 
In our case this driver run in a \texttt{Raspberry Pi3} with \texttt{raspbian OS} aboard the plane
\end{abstract}

\section{Introduction}

%MAVlink protocol

\section{UAV support in JdeRobot}

\section{MAVlink protocol}

\section{APM driver in JdeRobot}

\subsection{Design}
\label{sec:design}

The driver has to fullfil the following features:

\begin{itemize}
\item Have to be able to connect to a physical devices like APM 2.8 but as well as to a simutator.
\item Have to access to sensors measures of the aereal robot and interpreting the data and serve its as ICE interfaces objects.
\item Have to be able to recieve incoming commands as ICE interfaces objects and send its to the devices as \texttt{MAVLink's} commands.
\end{itemize}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{img/diseno.jpg}
  \caption{\texttt{APM Server} inputs and outputs}
  \label{fig:diseno_apms_caja_negra}
\end{figure}

In order to adress it, a Python in \texttt{JdeRobot} component has been developed who implements all of above features called \texttt{APM Server}. 
\texttt{APM Server} is organized in three logical layers.
\begin{itemize}
\item Comunication's layer with APM's devices. ensures the establishment and maintenance the comunication with the APM device through \texttt{MAVLink's} commands.
\item Bussiness Layer. In this layer the driver convert form \texttt{MAVLink's} commands to ICE interfaces objects and viceversa.
\item \texttt{JdeRobot comunication's layer} This layer create all the ICE servers needed and handle incoming and outgoing ICE interfaces objects.
\end{itemize}

\subsection{Comunication with APM devices layer}
\label{sec:apm_comunication}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head

In this layer the driver send the \texttt{MAVLink's} commands and recieve the \texttt{MAVLink's} messages which conform the comunication with the APM device.
The driver have to understand the APM device's messages received in \texttt{MAVLink} protocol as well as build this commands and send it to the APM.

\subsubsection{MAVLink's message}

The driver handle 11 commands of the full set of \texttt{MAVLink's} commands (see Figure 3.1) to allow the \texttt{JdeRobot} integration. Below it's described one of them in order to know this protocol.

\begin{verbatim}
<message id="33" name="GLOBAL_POSITION_INT">
	<description>The filtered global position 
	(e.g. fused GPS and accelerometers). The 
	position is in GPS-frame (right-handed, Z-up).
	It is designed as scaled integer message 
	since the resolution of float is not 
	sufficient.	</description>
<field type="uint32_t" name="time_boot_ms" units="ms">
	Timestamp (milliseconds since system boot)</field>
<field type="int32_t" name="lat" units="degE7">Latitude, 
	expressed as degrees * 1E7</field>
<field type="int32_t" name="lon" units="degE7">Longitude, 
	expressed as degrees * 1E7</field>
<field type="int32_t" name="alt" units="mm">Altitude in 
	meters, expressed as * 1000 (millimeters)</field>
<field type="int32_t" name="relative_alt" units="mm">
	Altitude above ground in meters, expressed as 
	* 1000 (millimeters)</field>
<field type="int16_t" name="vx" units="cm/s">Ground X 
	Speed (Latitude, positive north), expressed as 
	m/s * 100</field>
<field type="int16_t" name="vy" units="cm/s">Ground Y 
	Speed (Longitude, positive east), expressed as 
	m/s * 100</field>
<field type="int16_t" name="vz" units="cm/s">Ground Z 
	Speed (Altitude, positive down), expressed 
	as m/s * 100</field>
<field type="uint16_t" name="hdg" units="cdeg">Vehicle 
	heading (yaw angle) in degrees * 100, 0.0..359.99 
	degrees. If unknown, set to: UINT16_MAX</field>
</message>
\end{verbatim}

One example of this incoming message is:
\begin{verbatim}
GLOBAL_POSITION_INT {time_boot_ms : 480614, lat : -
353632612, lon : 1491652301, alt : 584110, 
relative_alt : -179, vx : 0, vy : 0, vz : 0, hdg : 
35608}
\end{verbatim}

\subsubsection{Setup and APM connection}

As has been showed the APM comunication is under \texttt{MAVLink} so the setup and conection have its own commands in this protocol.
To help us to build and understand these messages the driver uses \texttt{pymavlink}.

The connection is established in the server object creation, in the \_\_init\_\_() method of the class. The connection command needs two parameters: the device and the connection speed in bauds.
{\scriptsize
\begin{lstlisting}
classServer:
	def__init__(self,port,baudrate):
		self.master=mavutil.mavlink_connection(port,
					baudrate,autoreconnect=True)
		self.master.wait_heartbeat()
		...
__init__()
#test=Server("/dev/ttyUSB0",57600)#ConnectiontotherealAPMdevice
test=Server("udp:192.168.1.133:14558",57600)#ConnectiontoSITL
\end{lstlisting}}

The last two sentences shows an example of a real device connection and SITL's connection respectively.

Once the connection is made, the AMP \texttt{MAVLink's} messages start to be stored in a buffer, and we can setup it. The incoming messages rate and the set of these command we need are configurable parameters. Set up it is easy as is showed in the following lines.

{\scriptsize
\begin{lstlisting}
RATE=50
self.master.mav.request_data_stream_send(self.master.target_system,
						self.master.target_component,
						mavutil.mavlink.MAV_DATA_STREAM_ALL,
						RATE,1)
\end{lstlisting}}

The available set connections are:

\begin{verbatim}
MAV_DATA_STREAM_ALL	Enable all data streams
MAV_DATA_STREAM_RAW_SENSORS	Enable IMU_RAW, 
	        GPS_RAW, GPS_STATUS packets.
MAV_DATA_STREAM_EXTENDED_STATUS	Enable GPS_STATUS, 
	        CONTROL_STATUS, AUX_STATUS
MAV_DATA_STREAM_RC_CHANNELS	Enable RC_CHANNELS_SCALED,
	        RC_CHANNELS_RAW, SERVO_OUTPUT_RAW
MAV_DATA_STREAM_RAW_CONTROLLER	Enable 
	        ATTITUDE_CONTROLLER_OUTPUT, 
	        POSITION_CONTROLLER_OUTPUT, 
	        NAV_CONTROLLER_OUTPUT.
MAV_DATA_STREAM_POSITION	Enable LOCAL_POSITION, 
	        GLOBAL_POSITION/GLOBAL_POSITION_INT messages.
MAV_DATA_STREAM_EXTRA1	Dependent on the autopilot
MAV_DATA_STREAM_EXTRA2	Dependent on the autopilot
MAV_DATA_STREAM_EXTRA3	Dependent on the autopilot

\end{verbatim}

And the RATE value depends of the choosed device.

\subsubsection{APM data reading}
\label{subsec:apm_data_reading}

In order to retrieve the APM sensors measures, a thread is ejecuting a message handler all the time.
In this handler the method search for seven types of messages (like GLOBAL\_POSITION\_INT) which contains the needed data to fill the two ICE interfaces that the server serves: \texttt{Pose3D} y \texttt{NavData}.

\subsubsection{Sending missions to APM}
\label{sec:mission_apm}

This is the most important part of the driver. The driver recieve two ICE interfaces, one developed in this project called \texttt{mission} with the waypoints to follow and the \texttt{Extra} interface with the take off and land maneuvers.
Once a mission is received through these interfaces objects, \texttt{APM Server} build the \texttt{MAVLink's} commands needed to sent it to the APM. To perform it, a new thread with a listener was developed in order to know if there is a incominf mission.
Once this listener detects an incoming mission, call to \texttt{setMission(self, mission)} who build the \texttt{MAVLink's} commands and sent it as the Figure 3.3 diagram shows.
The \texttt{setMission(self, mission)} method read througth the list of waypoints and if \texttt{takeOffDecision} and \texttt{landDecision} of \texttt{Extra} object are seted to True add the \texttt{MAVLink's} commands of take off and land at first and last of the list of commans whitch is sended to the APM.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.70]{img/waypoint-protocol-sendlist.png}
  \caption{\texttt{MAVLink's} missions sending protocol}
  \label{fig:misiones_mavlink}
\end{figure}

\subsection{Connection to applications layer}
\label{sec:apm_jderobot_comunication}

In this layer all the ICE interfaces objects are handled. The driver uses four interfaces. Two of there are served (outgoing), \texttt{Pose3D} and \texttt{NavData} and the rest are received: \texttt{mission} and \texttt{extra}.
Each interfaces has its own thread with its own ICE server (note that ICE servers communication are bidirectional, so in this case ``server'' word don't refer to the side of the data).
In the following lines an ICE server implementation are showed.
{\scriptsize
\begin{lstlisting}

def openPose3DChannel(self, pose3D):
   status = 0
   ic = None
   Pose2Tx = pose3D
   try:
       ic = ICE.initialize(sys.argv)
       adapter = ic.createObjectAdapterWithEndpoints("Pose3DAdapter", 
       				"default -p 9998")
       object = Pose2Tx
       adapter.add(object, ic.stringToIdentity("ardrone_pose3d")) 
       adapter.activate()
       ic.waitForShutdown()
   except:
       traceback.print_exc()
       status = 1
   if ic:
       try:
           ic.destroy()
       except:
           traceback.print_exc()
           status = 1
   sys.exit(status)
\end{lstlisting}}

\subsection{Interpretation layer}
\label{interpretation_layer}

This layer makes all the needed data's tranformation. So once the \texttt{MAVLink's} message is located, this layer makes it fit in the corresponding ICE interfaces and the same in opposite side communication. 
The \texttt{setMission(self, mission)} and the \texttt{refreshAPMPose3D()} and
\texttt{refreshAPMnavdata()} are the most important methods involved in this layer.

\section{UAV Commander application}

\texttt{UAV Commander} is a JdeRobot application writen in python in order to offer to a human operator build and sent missions to an autonomous \texttt{MAVLink} plane.

\subsection{Design}
\label{design}

The application has to fulfill the following features:
\begin{itemize}
\item The application have to display all the sensors measures data.
\item Have to get a friendly end user interface what allow to build missions and follow its compliance.
\item Have to be able to send the builded mission to the drone.
\end{itemize}

\begin{figure}
\centering{
   \label{f:diseno_uavc_caja_negra}
    \includegraphics[width=0.8\textwidth]{img/diseno2_uavc.jpg}}
    \caption{\texttt{UAV Commander} input and outputs}

\end{figure}

The application is composed by 3 blocks: maps's block, missions's block and sensors's block. The maps's block is who adquire the maps and converts coordinates between diferent references systems. The missions's block allow to a human operator to build and sent missions and the sensors's block displays all de received sensors's measures data, onboad camera included.

\texttt{UAV Commander} connects to \texttt{APM Server} throught the  \texttt{Pose3D}, \texttt{NavData}, \texttt{Extra} and \texttt{mission} ICE interfaces, and can receive images and video throught \texttt{camera} from sources like JdeRobot's \texttt{cameraserver} driver.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.60]{img/arquitectura.jpg}
  \caption{\texttt{UAV Commander} and \texttt{APM Server} connection schema}
  \label{fig:uavc_ppal}
\end{figure}

\subsection{Maps's block}
\label{maps_block}

In this block the application download from Web Service Maps (WMS) a georeferenced map (the current version supports ``Instituto Geográfico Nacional de España'' (IGN) and Google as sources). For retrive these maps a few data is needed as: current possition (tatitude and longitude), the specific area to retrieve or bounding box, the desired image's resolution (in pixels) and the \textit{datum} or the Earth desired projection.
The WMS response is a map of the bounding box delimited zone. With this image and de input data its possible to infer the coordinates of a specific pixel in the image. In the current version of \texttt{UAV Commander} we agreed to despreciate the Earth curvature as long as we suppose the drone's missions local (less than 2 kilometers of radius) and these calculations are really simple.

\subsection{Missions's block}
\label{missions_block}

In this block the map is displayed and becomes interactive in order to help operator to build missions. If the maps is clicked in whatever position the application detects it, draw the waypoint mark in the map, infer the gographical possition and add that tho the waypoints table. And if a previous waypoint is created the application will draw the path from it to the new waypoint. Also is possible to make maneuvers like take off and land using the ``Take off/land'' button, pushing it and in a point of the map a take off is requested and the same in land maneuver.
In order to display the mission's tracking a complex layer system was developed, \texttt{UAV Commander} uses three layouts,  each one enriches the others. The first layer is the original downloaded image, is needed in case of a reset request, the second one comes frim the original too but is updated with a shadow in all the reached positions with the plane. And over this last layer the waypoints and the plane tracking and heading (the plane position are represented whit a triangle who rotates in order to reveal the heading of it) is drawed. This design allows resetting the map, clear the waypoints and not lost the shadows and tracking the plane with these all information.
In order to update all these information in the Graphical User Interface (GUI) a new thread was developed.
All these features are available from \texttt{UAV Commander} main window.

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{img/uavc.png}
  \caption{\texttt{UAV Commander} main window}
  \label{fig:uavc_ppal}
\end{figure}

\subsection{Sensors's block}
\label{sensors_block}

Here the sensors's measures data are displayed in way to be more understandables to the operator, in order to enrich the missions's tracking. This data is recovered from the APM Server in ICE interfaces objects, exactly Pose3D and NavData but has Camera ICE interface too so can receive images and video by this way, with JdeRobot's cameraserver driver for example. All these information are interpreted and displayed in the sensor's and camera's windows.

\begin{figure}
\centering{
   \label{f:actitud_uavc}
    \includegraphics[width=1\textwidth]{img/actitud.png}}
    \caption{\texttt{UAV Commander} attitude and battery load}

\end{figure}

\section{Conclusions}

\end{document}